### 2016-08-02 ch03-lrm/nonparametric.tex
Some Formatting
Some formatting + OE figures
Change spacing around headings, epigraphs, ...
Small changes for formatting
 - Moving labels in figures
 - Some negative spaces in formulas
Define T(P,C), adjust some figures
Stub: intro leakage
*_0 -> \true{*}

Here, we focus on using local modeling to separate the different terms, i.e.
\begin{itemize}
  \item the transfer function $\true{G}$,
  \item the leakage term $T$, and
  \item the noise term $V$.
\end{itemize}
The main term of interest, however, is the transfer function $\true{G}$ and the noise variance $\sigma^2_{V}$.

@@

\begin{remark}
Computing the inverse of this Hessian cannot, to the knowledge of the author, be reduced to a form that yields more insight.
However, if the non-diagonal blocks can be neglected, this would lead to an inverse that is also block-diagonal:
\begin{equation}
   \begin{bmatrix}
    \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_A \pd \theta_A } &
    \deemph{0} & 
    \deemph{0} \\
    \deemph{0} &
    \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_B \pd \theta_B } & 
    \deemph{0} \\
    \deemph{0} &
    \deemph{0} & 
    \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_C \pd \theta_C }
  \end{bmatrix}^{-1}
  \!\!\!=
  \begin{bmatrix}
    \left( \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_A \pd \theta_A }\right)^{-1} &
    \deemph{0} & 
    \deemph{0} \\
    \deemph{0} &
    \left( \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_B \pd \theta_B }\right)^{-1} & 
    \deemph{0} \\
    \deemph{0} &
    \deemph{0} & 
    \left( \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_C \pd \theta_C }\right)^{-1}
  \end{bmatrix}
  % \text{.}
\end{equation}
\end{remark}
\paragraph{Bias on the \glsentrydesc{FRF}}
As a first order approximation, the bias on the \gls{FRF} can be written as
\begin{equation}
  \Delta \LocalModel{G}(\omega_k) 
  =
  \Delta \LocalModel{G}(\dw{0}{k}) 
     \approx 
        \PartialDerivative{\LocalModel{G}(\dw{0}{k},\theta)}
                                          {\theta} 
      \Delta \theta
      \text{.}
\end{equation}
The derivative in that expression can be evaluated easily
\begin{align}
    \PartialDerivative{\LocalModel{G}(\dW, \theta)}{\theta} 
     &=
     \begin{bmatrix}
       \PartialDerivative{\LocalModel{G}(\dW,\theta)}{\theta_A} &
       \PartialDerivative{\LocalModel{G}(\dW,\theta)}{\theta_B} &
       \PartialDerivative{\LocalModel{G}(\dW,\theta)}{\theta_C} 
     \end{bmatrix}
     \\
     \PartialDerivative{\LocalModel{G}(\dW, \theta)}{\theta_A} 
     &=
     - \frac{\LocalModel{B}(\dW, \theta)}{\LocalModel{A}^2(\dW, \theta)}
     \begin{bmatrix}
         \dW^1 & \dW^2 & \deemph{\cdots} & \dW^{\order{A}}
     \end{bmatrix}
     \\
     \PartialDerivative{\LocalModel{G}(\dW, \theta)}{\theta_B} 
     &=
     \frac{1}{\LocalModel{A}(\dW, \theta)}
     \begin{bmatrix}
         1 & \dW^1 & \dW^2 & \deemph{\cdots} & \dW^{\order{B}}
     \end{bmatrix}
     \\
     \PartialDerivative{\LocalModel{G}(\dW, \theta)}{\theta_C} 
     &=
     \Zero{1 \times \order{C}+1}
     \text{.}
\end{align}

@@

\section{Smoothing the \glsentryshort{FRF} by Impulse Response Truncation}
\label{sec:nonparametric:truncation}
\researchBasedOn[This section]{Lumori2014TIM}

@@

Based on \citet{Schoukens2009LPM}, the nonparametric method developed in this work is formulated in an \gls{OE} framework. 
A \gls{SISO} \gls{LTI} system is considered.
The input signal is a known random noise signal and the output signal is corrupted by measurement noise. This is depicted in \figref{fig:nparam:blockH0} as a linear dynamic discrete-time \gls{SISO} system, whose full mathematical model is of the form:
\begin{equation}\label{eq:nparam:trunc:LPM:TD}
y(t)=\true{G}(q)u_0(t)+\true{H}(q)e(t)=\true{G}(q)u_0(t)+v(t)
\end{equation}
where $\true{G}(q)$ represents the dynamics of the system to be estimated, $u_0(t)$ is the input signal, $v(t)= \true{H}(q)e(t)$ is the noise source at the output, $\true{H}(q)$ is the noise dynamics, $e(t)$ is white noise, and $q^{-m}$ is the backward shift operator ($q^{-m}x(t)$ = $x(t-m)$  with $m$ a positive integer).

@@

These \gls{FRF} estimates are also compared with the true \gls{LTI} system ${G}_0(\omega_k)$.
Discussion of the simulation results is then followed by a conclusion in \secref{sec:nparam:trunc:conclusion}.

@@

% \begin{assumption}
% The measured output signal is $y(t) = y_0(t) + v(t)$ with the exact input signal $u_0(t)$ known. 
% The filtered white noise $v(t) = \true{H}(q)e(t)$ is the disturbing noise source at the output.
% \end{assumption}

@@

\end{enumerate}
This procedure is illustrated in \figref{fig:nparam:trunc:nonparametric:truncation:Ftest} for the system described by the following equations:
\begin{subequations}
\label{eq:trunc:systemODEs}
\begin{align}
y_0(t)  &= 1.5371y_0(t-1)    -0.9025y_0(t-2) + u(t)\\
y(t)       &= y_0(t) + e(t),
\end{align}
\end{subequations}
where $e(t)$ is a white noise sequence, such that the \gls{SNR} of the output signal is $78.2\unit{dB}$.
The figure shows segments of the estimated impulse response. The last two segments are dominated by the noise, while all samples at $t<80$ have large system contributions. The algorithm outlined above selects $\truncTime = 80$, beyond which all samples are set to zero, resulting in a smoothed \gls{FRF} estimate.

@@

\begin{equation}
y_0(t) = 2.583y_0(t - 1) -2.176y_0(t - 2)+0.592y_0(t-3) + u(t)
\end{equation}

The measured signal was disturbed by random white noise, $y(t) = y_0(t) + e(t)$, such that the \gls{SNR} was $14.2\unit{dB}$.
The result is depicted in \figref{fig:nparam:trunc:nonparametric:trunc:impresp:expfit}. 
An exponential function (black thick line) is fitted on the peaks of the absolute value of a noisy impulse response. 
The truncation time (vertical line) $\truncTime$ was selected as the time instant at which the fitted exponential fell below $0.4\noiseStd$ (i.e. $\gamma = 0.4$ in equation \eqref{eq:truncTimeExpFit}).

@@

\begin{figure}
    \centering
    \setlength\figurewidth{0.85\columnwidth}
    \setlength\figureheight{0.68\figurewidth}
    \input{\thisDir/figs/figLPMvsTRunc.tikz}
    \caption[Comparison of FRF estimated using LPM and Truncated LPM.]{The bias $\sampleBias{\placeholder}$ and standard deviation $\sampleStd{\placeholder}$ of the \gls{FRF} estimates using both the \gls{LPM} (blue) and truncated~\gls{LPM} (red) are shown. 
    It can be seen that the truncation reduces the variance at the cost of an increased bias, however, the overall \gls{RMS} error is reduced by the truncation.}
    \label{fig:nparam:trunc:LPMvsTRunc}
\end{figure}

@@

>>>>

Here, we focus on using local modeling to separate the different terms, i.e.
\begin{itemize}
  \item the transfer function $\true{G}$,
  \item the leakage term $T$, and
  \item the noise term $V$.
\end{itemize}
The main terms of interest are the transfer function $\true{G}$ and the noise variance $\sigma^2_{V}$.

@@

\begin{remark}
Computing the inverse of this Hessian cannot, to the knowledge of the author, be reduced to a form that yields more insight.
However, if the non-diagonal blocks can be neglected, this would lead to an inverse that is also block-diagonal:
\begin{equation}
   \begin{bmatrix}
    \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_A \pd \theta_A } \!\!\!\!\!\!&
    \deemph{0} & 
    \deemph{0} \\
    \deemph{0} &
    \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_B \pd \theta_B } \!\!\!\!\!\!& 
    \deemph{0} \\
    \deemph{0} &
    \deemph{0} & 
    \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_C \pd \theta_C } 
  \end{bmatrix}^{\!-1}
  \!\!=
  \begin{bmatrix}
    \left( \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_A \pd \theta_A }\right)^{\!-1} \!\!\!\!\!\!&
    \deemph{0} & 
    \deemph{0} \\
    \deemph{0} & \!\!\!\!\!\!
    \left( \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_B \pd \theta_B }\right)^{\!-1} \!\!\!\!\!\!& 
    \deemph{0} \\
    \deemph{0} &
    \deemph{0} & \!\!\!\!\!\!
    \left( \PartialDerivative[2]{\ELSCost{\theta}}{ \conj{\theta}_C \pd \theta_C }\right)^{\!-1} \!
  \end{bmatrix}
  % \text{.}
\end{equation}
\end{remark}
\paragraph{Bias on the \glsentrydesc{FRF}}
As a first order approximation, the bias on the \gls{FRF} can be written as
\begin{equation}
  \Delta \LocalModel{G}(\omega_k) 
  =
  \Delta \LocalModel{G}(\dw{0}{k}) 
     \approx 
        \PartialDerivative{\LocalModel{G}(\dw{0}{k},\theta)}
                                          {\theta} 
      \Delta \theta
      \text{.}
\end{equation}
The derivative in that expression can be evaluated easily
\begin{align}
    \PartialDerivative{\LocalModel{G}(\dW, \theta)}{\theta} 
     &=
     \begin{bmatrix}
       \PartialDerivative{\LocalModel{G}(\dW,\theta)}{\theta_A} &
       \PartialDerivative{\LocalModel{G}(\dW,\theta)}{\theta_B} &
       \PartialDerivative{\LocalModel{G}(\dW,\theta)}{\theta_C} 
     \end{bmatrix}
     \\
     \PartialDerivative{\LocalModel{G}(\dW, \theta)}{\theta_A} 
     &=
     - \frac{\LocalModel{B}(\dW, \theta)}{\LocalModel{A}^2(\dW, \theta)}
     \begin{bmatrix}
         \dW^1 & \dW^2 & \deemph{\cdots} & \dW^{\order{A}}
     \end{bmatrix}
     \\
     \PartialDerivative{\LocalModel{G}(\dW, \theta)}{\theta_B} 
     &=
     \frac{1}{\LocalModel{A}(\dW, \theta)}
     \begin{bmatrix}
         1 & \dW^1 & \dW^2 & \deemph{\cdots} & \dW^{\order{B}}
     \end{bmatrix}
     \\
     \PartialDerivative{\LocalModel{G}(\dW, \theta)}{\theta_C} 
     &=
     \Zero{1 \times \order{C}+1}
     \text{.}
\end{align}

@@

\section{Smoothing the \glsentryshort{FRF} by Impulse Response Truncation}
\label{sec:nonparametric:truncation}
\sectionBasedOn{Lumori2014TIM}

@@

Based on \citet{Schoukens2009LPM}, the nonparametric method developed in this work is formulated in an \gls{OE} framework. 
A \gls{SISO} \gls{LTI} system is considered.
The input signal is a known random noise signal and the output signal is corrupted by measurement noise. This is depicted in \figref{fig:nparam:blockH0} as a linear dynamic discrete-time \gls{SISO} system, whose full mathematical model is of the form:
\begin{equation}\label{eq:nparam:trunc:LPM:TD}
y(t)=\true{G}(q)\true{u}(t)+\true{H}(q)e(t)=\true{G}(q)\true{u}(t)+v(t)
\end{equation}
where $\true{G}(q)$ represents the dynamics of the system to be estimated, $\true{u}(t)$ is the input signal, $v(t)= \true{H}(q)e(t)$ is the noise source at the output, $\true{H}(q)$ is the noise dynamics, $e(t)$ is white noise, and $q^{-m}$ is the backward shift operator ($q^{-m}x(t)$ = $x(t-m)$  with $m$ a positive integer).

@@

These \gls{FRF} estimates are also compared with the true \gls{LTI} system $\true{G}(\omega_k)$.
Discussion of the simulation results is then followed by a conclusion in \secref{sec:nparam:trunc:conclusion}.

@@

% \begin{assumption}
% The measured output signal is $y(t) = \true{y}(t) + v(t)$ with the exact input signal $\true{u}(t)$ known. 
% The filtered white noise $v(t) = \true{H}(q)e(t)$ is the disturbing noise source at the output.
% \end{assumption}

@@

\end{enumerate}
This procedure is illustrated in \figref{fig:nparam:trunc:nonparametric:truncation:Ftest} for the system described by the following equations:
\begin{subequations}
\label{eq:trunc:systemODEs}
\begin{align}
\true{y}(t)  &= 1.5371\true{y}(t-1)    -0.9025 \true{y}(t-2) + u(t)\\
y(t)       &= \true{y}(t) + e(t),
\end{align}
\end{subequations}
where $e(t)$ is a white noise sequence, such that the \gls{SNR} of the output signal is $78.2\unit{dB}$.
The figure shows segments of the estimated impulse response. The last two segments are dominated by the noise, while all samples at $t<80$ have large system contributions. The algorithm outlined above selects $\truncTime = 80$, beyond which all samples are set to zero, resulting in a smoothed \gls{FRF} estimate.

@@

\begin{equation}
\true{y}(t) = 2.583\true{y}(t - 1) -2.176\true{y}(t - 2)+0.592\true{y}(t-3) + u(t)
\end{equation}

The measured signal was disturbed by random white noise, $y(t) = \true{y}(t) + e(t)$, such that the \gls{SNR} was $14.2\unit{dB}$.
The result is depicted in \figref{fig:nparam:trunc:nonparametric:trunc:impresp:expfit}. 
An exponential function (black thick line) is fitted on the peaks of the absolute value of a noisy impulse response. 
The truncation time (vertical line) $\truncTime$ was selected as the time instant at which the fitted exponential fell below $0.4\noiseStd$ (i.e. $\gamma = 0.4$ in equation \eqref{eq:truncTimeExpFit}).

@@

\begin{figure}
    \centering
    \setlength\figurewidth{0.85\columnwidth}
    \setlength\figureheight{0.62\figurewidth}
    \input{\thisDir/figs/figLPMvsTRunc.tikz}
    \caption[Comparison of FRF estimated using LPM and Truncated LPM.]{The bias $\sampleBias{\placeholder}$ and standard deviation $\sampleStd{\placeholder}$ of the \gls{FRF} estimates using both the \gls{LPM} (blue) and truncated~\gls{LPM} (red) are shown. 
    It can be seen that the truncation reduces the variance at the cost of an increased bias, however, the overall \gls{RMS} error is reduced by the truncation.}
    \label{fig:nparam:trunc:LPMvsTRunc}
\end{figure}

@@

\begin{subappendices}
  \input{\thisDir/leakage.tex}
\end{subappendices}

### 2016-08-02 ch05-initvals/initial-values.tex
Some minor formatting things
Some formatting + OE figures
Small changes for formatting
 - Moving labels in figures
 - Some negative spaces in formulas
*_0 -> \true{*}

\chapter{Initialization of Parametric Estimators}
\label{sec:initvals}
\myEpigraph{Everything starts somewhere, though many physicists disagree. But people have always been dimly aware of the problem with the start of things. They wonder how the snowplough driver gets to work, or how the makers of dictionaries look up the spelling of words.}{Terry Pratchett}{Hogfather}
\gdef\thisDir{ch05-initvals}
\tikzsetfigurename{ch05fig}
\glsresetall
\researchBasedOn{Geerardyn2015TIM}

@@

The \gls{RFIR} estimator is formulated in the time domain. 
It estimates the impulse response of a discrete time system as the minimizer of the following regularized least squares objective function:
\begin{align}\label{eq:hRFIRdef}
\hat g_{\RFIR} &= \argmin_{g} \norm[2]{y - g \conv u }^2 + \sigma^2 g^{\TT}P^{-1}g
\end{align}
where $g$ is the vectorized impulse response $g(t)$, with $t = 0,1,\cdots,n_h - 1$, assuming that the impulse response is $n_g$ samples long. 
Furthermore, $g\conv u$ is the convolution of $g$ with the input signal $u$, and $P\in \mathbb{R}^{n_g\times n_g}$ is the kernel matrix~\citep{Chen2012}. 
Note that \eqref{eq:hRFIRdef} is quadratic in $g$, such that $\hat g_{\RFIR}$ can be computed analytically in a single step.
The kernel matrix $P$ embodies prior knowledge on the system to be estimated. 
Here, the \gls{DCkernel}~\citep{Chen2012,Pillonetto2010} is used. 
Specifically, the element at $t_1,t_2$ of $P$ is given by
\begin{equation}
P(\alpha,\beta)_{t_1,t_2} = e^{-\alpha|t_1 - t_2|}e^{-\beta(t_1 + t_2)/2}
\end{equation}
 for $t_1,t_2 \in \set{0,1,\cdots,n_h - 1}$ which makes the estimated impulse response to be an exponentially decaying function of $t$ with decay rate $\beta^{-1}$, and correlation length $\alpha^{-1}$. 
 The latter \emph{hyperparameters} $\alpha$, $\beta$ and $\sigma$ are determined using Empirical Bayes~\citep{Carlin2000,Gelman2014}, i.e. as the maximizers of the \gls{LML} of the measured output signal~\citep{Chen2012}:
\begin{align}
\mathrm{LML}(y) = -y^{\TT}\Sigma(\alpha,\beta,\sigma)^{-1}y - \log |\Sigma(\alpha,\beta,\sigma)| \text{,}
\end{align}
where $\Sigma(\alpha,\beta, \sigma) = \sigma^2 \I + \phi^{\TT}P(\alpha,\beta)\phi$ and $\phi$ is a Toeplitz matrix constructed with $u$. 
This problem is solved using \mcode{fmincon} in \MATLAB.
%TODO: check convexity of this problem

@@

\begin{figure}
  \centering
  \setlength{\figurewidth}{0.8\columnwidth}
  \setlength{\figureheight}{0.6\figurewidth}
  \input{\thisDir/figs/MC-SNR1.1dB-NBW36-Q6dB.tikz}
 \caption[Model error for different initialization schemes on simulations.]{The distance $\|\model{\bullet} - \hat{G}_0 \|_2$  between the different estimates ($\model{exist}$~\legref{leg:dist-G-exist}, $\model{trunc}$~\legref{leg:dist-G-trunc} and $\model{RFIR}$~\legref{leg:dist-G-RFIR} ) and the `best theoretical' estimates ($\model{0}$) shows that the smoothers help the estimate to converge to $\model{0}$.
 The same measure is shown for the initial values $\model[init]{RFIR}$ \legref{leg:dist-G-RFIR-init} and $\model[init]{trunc}$ \legref{leg:dist-G-trunc-init}.
 $\|{G}_{0} - \hat{G}_0\|_2$\legref{leg:dist-G0} shows how far the true model and `best theoretical' estimates are apart to put the other distances into perspective. }
  \label{fig:single-sim}
\end{figure}

@@

%TODO: TM: maybe look into power consumption aspects?

>>>>

\chapter{Initialization of Parametric Estimators}
\label{sec:initvals}
\chapterBasedOn{Geerardyn2015TIM}
\myEpigraph{Everything starts somewhere, though many physicists disagree. But people have always been dimly aware of the problem with the start of things. They wonder how the snowplough driver gets to work, or how the makers of dictionaries look up the spelling of words.}{Terry Pratchett}{Hogfather}
\gdef\thisDir{ch05-initvals}
\tikzsetfigurename{ch05fig}
\glsresetall{}

@@

The \gls{RFIR} estimator is formulated in the time domain. 
It estimates the impulse response of a discrete time system as the minimizer of the following regularized least squares objective function:
\begin{align}\label{eq:hRFIRdef}
\hat g_{\RFIR} &= \argmin_{g} \norm[2]{y - g \conv u }^2 + \sigma^2 g^{\TT}P^{-1}g
\end{align}
where $g$ is the vectorized impulse response $g(t)$, with $t = 0,1,\cdots,n_h - 1$, assuming that the impulse response is $n_g$ samples long. 
Furthermore, $g\conv u$ is the convolution of $g$ with the input signal $u$, and $P\in \mathbb{R}^{n_g\times n_g}$ is the kernel matrix~\citep{Chen2012}. 
Note that \eqref{eq:hRFIRdef} is quadratic in $g$, such that $\hat g_{\RFIR}$ can be computed analytically in a single step.
The kernel matrix $P$ embodies prior knowledge on the system to be estimated. 
Here, the \gls{DCkernel}~\citep{Chen2012,Pillonetto2010} is used. 
Specifically, the element at $t_1,t_2$ of $P$ is given by
\begin{equation}
P(\alpha,\beta)_{t_1,t_2} = e^{-\alpha|t_1 - t_2|}e^{-\beta(t_1 + t_2)/2}
\end{equation}
 for $t_1,t_2 \in \set{0,1,\cdots,n_h - 1}$ which makes the estimated impulse response to be an exponentially decaying function of $t$ with decay rate $\beta^{-1}$, and correlation length $\alpha^{-1}$. 
 The latter \emph{hyperparameters} $\alpha$, $\beta$ and $\sigma$ are determined using Empirical Bayes~\citep{Carlin2000,Gelman2014}, i.e. as the maximizers of the \gls{LML} of the measured output signal~\citep{Chen2012}:
\begin{align}
\mathrm{LML}(y) = -y^{\TT}\Sigma(\alpha,\beta,\sigma)^{-1}y - \log |\Sigma(\alpha,\beta,\sigma)| \text{,}
\end{align}
where $\Sigma(\alpha,\beta, \sigma) = \sigma^2 \I + \phi^{\TT}P(\alpha,\beta)\phi$ and $\phi$ is a Toeplitz matrix constructed with $u$. 
This problem is solved using \mcode{fmincon} in \MATLAB.

@@

\begin{figure}
  \centering
  \setlength{\figurewidth}{0.8\columnwidth}
  \setlength{\figureheight}{0.6\figurewidth}
  \input{\thisDir/figs/MC-SNR1.1dB-NBW36-Q6dB.tikz}
 \caption[Model error for different initialization schemes on simulations.]{The distance $\|\model{\bullet} - \model{\true{}} \|_2$  between the different estimates ($\model{exist}$~\legref{leg:dist-G-exist}, $\model{trunc}$~\legref{leg:dist-G-trunc} and $\model{RFIR}$~\legref{leg:dist-G-RFIR} ) and the `best theoretical' estimates ($\model{\true{}}$) shows that the smoothers help the estimate to converge to $\model{\true{}}$.
 The same measure is shown for the initial values $\model[init]{RFIR}$ \legref{leg:dist-G-RFIR-init} and $\model[init]{trunc}$ \legref{leg:dist-G-trunc-init}.
 $\norm[2]{\true{G} - \model{\true{}}}$\legref{leg:dist-G0} shows how far the true model and `best theoretical' estimates are apart to put the other distances into perspective. }
  \label{fig:single-sim}
\end{figure}

@@
